---
title: "caseStudy2-version2"
author: "Gloria"
date: "2017/11/1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

response: specific time; 
```{r}
library(xlsx)
setwd("C:/Users/Lenovo/Desktop/STAT 440/caseStudy 2")
dat <- read.csv("Seattle_Police_Department_911_Incident_Response.csv", header = TRUE, sep = ",")
dat1 <- dat # dat1 is the backup to save time, if screw up somehow, use dat<-dat1 to recover
dat <- dat[c(4,5,7,8,11,13:14)]
# [1] "Event.Clearance.Code" "Event.Clearance.Date" "Zone.Beat" "Longitude"           
# [5] "Latitude"   

levels(dat$Event.Clearance.Code)
dat$Event.Clearance.Code <- as.character(dat$Event.Clearance.Code)
dat <- dat[dat$Event.Clearance.Code!='',]
dat <- dat[dat$Event.Clearance.Code!='NULL',]
dim(dat)
# double check
levels(as.factor(dat$Event.Clearance.Code))

levels(dat$Zone.Beat)
dat$Zone.Beat <- as.character(dat$Zone.Beat)
dat <- dat[dat$Zone.Beat != '',]
dim(dat)
# double check
levels(as.factor(dat$Zone.Beat))

sort(prop.table(table(dat$Event.Clearance.Code)), decreasing = TRUE)[1:5]
sort(table(dat$Event.Clearance.Code), decreasing = FALSE)

levels(dat$Event.Clearance.Date)
dat$Event.Clearance.Date <- as.character(dat$Event.Clearance.Date)
dat <- dat[dat$Event.Clearance.Date != '',]
dim(dat)

levels(as.factor(dat$Longitude))
levels(as.factor(dat$Latitude))
range(dat$Longitude)
range(dat$Latitude)


code <- read.csv(url("https://raw.githubusercontent.com/jasonjang714/911_Call_for_Services_Case_Study/master/seattle_code.csv"), as.is = TRUE, na.strings = c("",NA))
code1 <- code
code <- code[which(!is.na(code$MIR.Codes)),]
event_clearance_code_investg <- code$MIR.Codes
dat2 <- dat[dat$Event.Clearance.Code %in% event_clearance_code_investg,]
sort(prop.table(table(dat2$Event.Clearance.Code)), decreasing = TRUE)[1:5]
sort(table(dat2$Event.Clearance.Code), decreasing = TRUE)[1:5]
write.csv(code,"seattle_code.csv")


```


GOAL:Can you build a method to suggest what times of day, week, or year that a greater police presence would be useful
```{r}
Sys.setenv("LANGUAGE"="En")
Sys.setlocale("LC_ALL", "English")
library(dplyr)
library(stringr)
dsm <- dat2
date_time <- as.data.frame(str_split_fixed(dsm$Event.Clearance.Date, " ",3),stringsAsFactors = FALSE)
names(date_time) <- c("MM.DD.YY","spc_time","AM.PM")

MM.DD.YY <- as.data.frame(date_time$MM.DD.YY)
spc_time <- as.data.frame(date_time$spc_time)
AM.PM <- as.data.frame(date_time$AM.PM)

colnames(MM.DD.YY) <- 'MM.DD.YY'
colnames(spc_time) <- 'spc_time'
colnames(AM.PM) <- 'AM.PM'

dsm <- cbind(dsm,MM.DD.YY, spc_time, AM.PM)

```

Explainatory Analysis: histograms of date
```{r}
split_mdy <- as.data.frame((str_split_fixed(MM.DD.YY$MM.DD.YY, "/", 3)), stringsAsFactors = FALSE)
mm <- as.data.frame(as.numeric(unlist(split_mdy[1])))
dd <- as.data.frame(as.numeric(unlist(split_mdy[2])))
yy <- as.data.frame(as.numeric(unlist(split_mdy[3])))
hist(mm[,1])
hist(dd[,1])
hist(yy[,1])
# how about a week?
MM.DD.YY$MM.DD.YY <- strptime(MM.DD.YY$MM.DD.YY,format="%m/%d/%Y")
barplot(prop.table(sort(table(weekdays(MM.DD.YY$MM.DD.YY)))))

```

Explainatory Analysis: histogram of times
```{r}
require(data.table)
timecmb <- cbind(spc_time,AM.PM)
timecmb$cmd <- paste(timecmb$spc_time, timecmb$AM.PM, sep=" ")
timecmb$spc_time <- NULL
timecmb$AM.PM <- NULL

timecmb$cmd <- strptime(timecmb[,1], format="%I:%M:%S %p")
timecmb <- as.data.frame(lapply(timecmb,as.character), stringsAsFactors = FALSE)
func <- function(x)
{
  x <- str_split_fixed(as.character(x)," ",2)[2]
  return(x)
}
timecmb <- transpose(as.data.frame(lapply(timecmb$cmd, func), stringsAsFactors = FALSE))

split_hms <- as.data.frame((str_split_fixed(timecmb$V1, ":", 3)), stringsAsFactors = FALSE)
hh <- as.data.frame(as.numeric(unlist(split_hms[1])))
mminute <- as.data.frame(as.numeric(unlist(split_hms[2])))
ss <- as.data.frame(as.numeric(unlist(split_hms[3])))
names(hh) <- "hh"
names(mminute) <- "mminute"
names(ss) <- "ss"

hist(hh[,1])
hist(mminute[,1])
hist(ss[,1])
```

Explainatory Analysis: analyze the whole year
```{r}
require(plyr)
sum_date <- cbind(mm,dd,yy,timecmb)
names(sum_date) <- c("mm","dd","yy","timecmb")
sum_date$yy <- 2017
sum_date$mmddyy <- paste(sum_date$mm,sum_date$dd,sum_date$yy,sep='/')
sum_date$sum_date <- paste(sum_date$mmddyy,sum_date$timecmb,sep=" ")
# sum_date <- as.data.frame(lapply(sum_date, as.factor))
sum_date <- as.data.frame(sum_date$sum_date, stringsAsFactors = FALSE)
sum_date <- na.omit(sum_date)
names(sum_date) <- "sum_date1"
require(plyr)
sum_date_count_table <- count(sum_date$sum_date1) # weird... used to be count(sum_date$sum_date1)
more_times_incidents <-sum_date_count_table[which(sum_date_count_table$n >= 8),]
less_times_incidents <- sum_date_count_table[which(sum_date_count_table$n < 8),]
```

switch MM.DD.YY to weekdays - response
set AM.PM as factor variables
some MM.DD.YY not in standard form (i.e, cannot change to date)
```{r}
lct <- Sys.getlocale("LC_TIME")
Sys.setlocale("LC_TIME", "C")
Sys.setlocale("LC_TIME", lct)
weekday <- weekdays(as.Date(dsm$MM.DD.YY, "%m/%d/%Y"))
weekday <- as.factor(weekday)
ds_full <- cbind(weekday, dsm)
sum_day <- NA
ds_full <- cbind(sum_day, ds_full, hh, mminute, ss)
ds_full$Event.Clearance.Group <- as.factor(ds_full$Event.Clearance.Group)



night <- c(0:5)
morning <- c(6:12)
noon <- c(13:17)
evening <- c(18:24)
ds_full$sum_day[which(ds_full$hh %in% morning)] <- "morning"
ds_full$sum_day[which(ds_full$hh %in% noon)] <- "noon"
ds_full$sum_day[which(ds_full$hh %in% evening)] <- "evening"
ds_full$sum_day[which(ds_full$hh %in% night)] <- "night"



ds_full$sum_day <- as.factor(ds_full$sum_day)
ds_full$Event.Clearance.Description <- as.factor(ds_full$Event.Clearance.Description)
names(ds_full)
ds_full <- na.omit(ds_full)


ds_full1 <- ds_full[, -c(4:6, 10:11, 13:15)]
ds_full1$Event.Clearance.Code <- as.factor(ds_full$Event.Clearance.Code)
ds_full1$Zone.Beat <- as.factor(ds_full$Zone.Beat)


write.csv(ds_full1, "Seattle_0.005_analysis.csv") 
```

LASSO model - sum_day as response
```{r}
library(glmnet)
library(nnet)
set.seed(1234)

## Possibility 1: Take a sample of 0.5% of the rows
fraction = 0.005
set.seed(12345) ## Change this to your own number
sample_rows = sample(1:nrow(ds_full1), floor(nrow(ds_full1) * fraction))
ds_full2 = ds_full1[sample_rows,]
#ds_full2 <- subset(ds_full2, select = -c(hh,mm))

train <- ds_full2[1:round(0.8*nrow(ds_full2)), ]
test <- ds_full2[round(0.8*nrow(ds_full2)+1):nrow(ds_full2), -1]
y_real <- as.vector(ds_full2[round(0.8*nrow(ds_full2)+1):nrow(ds_full2), 1])

x_lasso <- data.matrix(train[,-1])
y_lasso <- train[,1]
fit_lasso <- cv.glmnet(x_lasso, y_lasso, family="multinomial", type.measure = "class")

fit_lasso

fit_lasso$lambda.min
fit_lasso$lambda.1se
plot(fit_lasso)


coef(fit_lasso, s="lambda.min")
coef(fit_lasso, s="lambda.1se")

newX <- data.matrix(test)

pred_lasso_min <- predict(fit_lasso, newx=newX, type = "class", s="lambda.min")
result_lasso_min <- mapply(grepl, pred_lasso_min, y_real)
accRate_lasso_min <- sum(result_lasso_min)/length(pred_lasso_min)
accRate_lasso_min
```
Confusion matrix
introduction: https://cran.r-project.org/web/packages/caret/caret.pdf
```{r}
library(caret)
library(e1071)
confusionMatrix(pred_lasso_min, y_real)
```

Logit - sum_Day as response
```{r}
fit_log <- multinom(sum_day~., data = train)
pred_log <- predict(fit_log, newdata = test, "class")
compare_log <- cbind(y_real, as.numeric(pred_log))
colnames(compare_log) <- c("real value", "predict value")
compare_log <- as.data.frame(compare_log)
compare_log$`real value` <- as.numeric(compare_log$`real value`)
compare_log$`predict value` <- as.numeric(compare_log$`predict value`)

log_result1 <- 0
for (i in 1:nrow(compare_log)){
  if (compare_log$`real value`[i] == compare_log$`predict value`[i]){
    log_result1 <- log_result1+1
  }
  else{
    log_result1 <- log_result1
  }
}

accRate_log1 <- log_result1/nrow(compare_log)
accRate_log1
```
Confusion matrix for logit model
```{r}
confusionMatrix(pred_log, y_real)
```






