---
title: "CaseStudy 2"
author: "Gloria"
date: "2017/10/20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(xlsx)
setwd("C:/Users/Lenovo/Desktop/STAT 440/caseStudy 2")
dat <- read.csv("Seattle_Police_Department_911_Incident_Response.csv", header = TRUE, sep = ",")
names(dat)[1] <- "CAD.CDW.ID"
fraction <- 0.05
set.seed(1234)
sample_rows <- sample(1:nrow(dat), floor(nrow(dat)*fraction))

dat5 <- dat[sample_rows,]
write.csv(dat5, "Seattle_0.05.csv")


dat = dat5 # overwrite the old dat
dat5 = NULL # To clear RAM
gc()
```

GOAL:Can you build a method to suggest what times of day, week, or year that a greater police presence would be useful
```{r}
Sys.setenv("LANGUAGE"="En")
Sys.setlocale("LC_ALL", "English")
ds <- read.csv("Seattle_0.05.csv", header = TRUE, sep=",")
ds <- ds[,-1]
head(ds)
names(ds)
```


```{r}
library(dplyr)
dsm <- ds[c(1,2,4,5,7,8,13,14)]
#"CAD.CDW.ID",  "CAD.Event.Number" ,"Event.Clearance.Code" 
#"Event.Clearance.Description" , "Event.Clearance.Group" ,
#"Event.Clearance.Date" , "Longitude" , "Latitude"
dsm <- subset(dsm, dsm$Event.Clearance.Group == "AUTO THEFTS")
head(dsm)
for(i in 1:8){
  if(sum(is.na(dsm[,i]))==0){
    print("CHECK")
  }
  else{
    print("NA")
  }
}

library(stringr)
dsm[,6] <- as.character(dsm[,6])
date_time <- as.data.frame(str_split_fixed(dsm[,6], " ",3),stringsAsFactors = FALSE)
MM.DD.YY <- as.data.frame(date_time[,1], stringsAsFactors = FALSE)

spc_time <- as.data.frame(date_time[,2], stringsAsFactors = F)
AM.PM <- as.factor(date_time[,3])
names(MM.DD.YY) <- c("MM.DD.YY")
names(spc_time) <- c("spc_time")
names(AM.PM) <- c("AM.PM")

dsm <- cbind(dsm,MM.DD.YY, spc_time, AM.PM)

```

Explainatory Analysis: histograms of date
```{r}
split_mdy <- as.data.frame((str_split_fixed(MM.DD.YY$MM.DD.YY, "/", 3)), stringsAsFactors = FALSE)
mm <- as.data.frame(as.numeric(unlist(split_mdy[1])))
dd <- as.data.frame(as.numeric(unlist(split_mdy[2])))
yy <- as.data.frame(as.numeric(unlist(split_mdy[3])))
hist(mm[,1])
hist(dd[,1])
hist(yy[,1])
# how about a week?
MM.DD.YY$MM.DD.YY <- strptime(MM.DD.YY$MM.DD.YY,format="%m/%d/%Y")
barplot(prop.table(table(weekdays(MM.DD.YY$MM.DD.YY))))
```

Explainatory Analysis: histogram of times /// error in transpose function
```{r}
require(data.table)
timecmb <- cbind(spc_time,AM.PM)
timecmb$cmd <- paste(timecmb$spc_time, timecmb$AM.PM, sep=" ")
timecmb$spc_time <- NULL
timecmb$AM.PM <- NULL

timecmb$cmd <- strptime(timecmb[,1], format="%I:%M:%S %p")
timecmb <- as.data.frame(lapply(timecmb,as.character), stringsAsFactors = FALSE)
func <- function(x)
{
  x <- str_split_fixed(as.character(x)," ",2)[2]
  return(x)
}
timecmb <- transpose(as.data.frame(lapply(timecmb$cmd, func), stringsAsFactors = FALSE))

split_hms <- as.data.frame((str_split_fixed(timecmb$V1, ":", 3)), stringsAsFactors = FALSE)
hh <- as.data.frame(as.numeric(unlist(split_hms[1])))
mminute <- as.data.frame(as.numeric(unlist(split_hms[2])))
ss <- as.data.frame(as.numeric(unlist(split_hms[3])))
hist(hh[,1])
hist(mminute[,1])
hist(ss[,1])
```

Explainatory Analysis: analyze the whole year
```{r}
require(plyr)
sum_date <- cbind(mm,dd,yy,timecmb)
names(sum_date) <- c("mm","dd","yy","timecmb")
sum_date$yy <- 2017
sum_date$mmddyy <- paste(sum_date$mm,sum_date$dd,sum_date$yy,sep='/')
sum_date$sum_date <- paste(sum_date$mmddyy,sum_date$timecmb,sep=" ")
sum_date <- as.data.frame(lapply(sum_date, as.factor))
sum_date <- as.data.frame(sum_date$sum_date, stringsAsFactors = FALSE)
sum_date <- na.omit(sum_date)
names(sum_date) <- "sum_date1"
sum_date_count_table <- count(sum_date$sum_date1)
three_times_incidents <-sum_date_count_table[which(sum_date_count_table$freq > 2),]
two_times_incidents <- sum_date_count_table[which(sum_date_count_table$freq == 2),]
```

Clearning: count() is a better version of table()
1.Event.Clearance.Code is related to the Event.Clearance.Desscription
2. rare data cleaning
```{r}
dsm$Event.Clearance.Description <- as.character(dsm$Event.Clearance.Description)
require(plyr)
des_count <- count(dsm$Event.Clearance.Description)
group_count <- count(dsm$Event.Clearance.Group)

rare_des <- des_count[des_count$freq <= 5,]
rare_des_x <- as.vector(rare_des$x)

dsm <- subset(dsm,  ! dsm$Event.Clearance.Description %in% rare_des_x)

rare_group <- group_count[group_count$freq <= 5,]
rare_group_x <- as.vector(rare_group$x)
dsm <- subset(dsm, ! dsm$Event.Clearance.Group %in% rare_group_x)


head(dsm)
```


1. switch MM.DD.YY to weekdays - response
2. set AM.PM as factor variables
3. some MM.DD.YY not in standard form (i.e, cannot change to date)
```{r}
lct <- Sys.getlocale("LC_TIME")
Sys.setlocale("LC_TIME", "C")
Sys.setlocale("LC_TIME", lct)
weekday <- weekdays(as.Date(dsm$MM.DD.YY, "%m/%d/%Y"))
weekday <- as.factor(weekday)
ds_full <- cbind(weekday, dsm)
ds_full <- na.omit(ds_full)
ds_full$Event.Clearance.Group <- as.factor(ds_full$Event.Clearance.Description)

day <- as.data.frame(str_split_fixed(ds_full$spc_time, ":", 3), stringAsFactor=F)
hr <- day[,1]
mins <-day[,2]
sec <- day[,3]
sum_day <- NA
ds_full <- cbind(sum_day, ds_full, hr, mins, sec)


for (i in 1:nrow(ds_full)){

 if (ds_full$AM.PM[i] =="AM"){
    if (ds_full$hr[i] %in% c("00","01","02","03","04","05")){
      ds_full$sum_day[i] <- "night"
    }
    else{
      ds_full$sum_day[i] <-"morning"
    }
  }
  if (ds_full$AM.PM[i] == "PM"){
    if (ds_full$hr[i] %in% c("00","01","02","03","04","05")){
     ds_full$sum_day[i] <-"noon"
    }
    else{
      ds_full$sum_day[i] <-"evening"
    }
  }
}
ds_full$sum_day <- as.factor(ds_full$sum_day)
ds_full$Event.Clearance.Description <- as.factor(ds_full$Event.Clearance.Description)
names(ds_full)
ds_full <- subset(ds_full, select = -c(CAD.CDW.ID, CAD.Event.Number,Event.Clearance.Date,Event.Clearance.Group, Event.Clearance.Code, MM.DD.YY, spc_time, hr, mins, sec))
write.csv(ds_full, "Seattle_0.05_analysis.csv")
```

Regression model
LASSO
```{r}
library(glmnet)
library(nnet)
set.seed(1234)
train <- ds_full[1:round(0.85*nrow(ds_full)), ]
test <- ds_full[round(0.85*nrow(ds_full)+1):nrow(ds_full), -1]
y_real <- as.vector(ds_full[round(0.85*nrow(ds_full)+1):nrow(ds_full), 1])

x_lasso <- data.matrix(train[,-1])
y_lasso <- train[,1]
fit_lasso <- cv.glmnet(x_lasso, y_lasso, family="multinomial", type.measure = "class")
fit_lasso

fit_lasso$lambda.min
fit_lasso$lambda.1se
plot(fit_lasso)


coef(fit_lasso, s="lambda.min")
coef(fit_lasso, s="lambda.1se")

newX <- data.matrix(test)

pred_lasso_min <- predict(fit_lasso, newx=newX, type = "class", s="lambda.min")
result_lasso_min <- mapply(grepl, pred_lasso_min, y_real)
accRate_lasso_min <- sum(result_lasso_min)/length(pred_lasso_min)
accRate_lasso_min
```

logit
```{r}
fit_log <- multinom(sum_day~., data = train)
pred_log <- predict(fit_log, newdata = test, "class")
compare_log <- cbind(y_real, as.numeric(pred_log))
colnames(compare_log) <- c("real value", "predict value")
compare_log

```


regression tree
```{r}
library(boot)
library(rpart)
fit_rtree <- rpart(sum_day~.,data=train, method="class")
plot(fit_rtree, uniform = T)
text(fit_rtree, use.n = T, all = T, cex=0.5)
```






Examine the data
```{r}
library(lubridate)
str(ds)
ds$CAD.CDW.ID <- as.factor(ds$CAD.CDW.ID)
ds$CAD.Event.Number <- as.factor(ds$CAD.Event.Number)
ds$General.Offense.Number <- as.factor(ds$General.Offense.Number)
ds$Event.Clearance.Date <- as.character(mdy_hms(ds$Event.Clearance.Date, tz=''))

ds$At.Scene.Time <- as.Date(ds$At.Scene.Time)
for (i in 1:ncol(ds))
{
  print(sprintf("%s: %s",colnames(ds)[i],nrow(count(ds[,i]))))
}
count(ds$District.Sector) # NULL:22
count(ds$Zone.Beat)
count(ds$Initial.Type.Description)
count(ds$Event.Clearance.Group) # NULL:51
count(ds$Initial.Type.Subgroup)
```

